---
title: "Activity Monitoring Data Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(knitr)
library(ggplot2)
knitr::opts_chunk$set(fig.path="figure/", echo=TRUE, message=FALSE, warning=FALSE)
```

## Loading and Preprocessing the Data

```{r load-data}
# Read the data
activity_data <- read.csv("activity.csv", na.strings = "NA")

# Convert date column to Date type
activity_data$date <- as.Date(activity_data$date, format="%Y-%m-%d")
```

## Total Number of Steps Taken Per Day

```{r total-steps}
# Calculate total steps per day
steps_per_day <- aggregate(steps ~ date, activity_data, sum, na.rm = TRUE)

# Plot histogram
hist(steps_per_day$steps, breaks=20, col="blue", main="Histogram of Total Steps Per Day", xlab="Total Steps")

# Save figure
png("figure/total_steps_hist.png", width=800, height=600)
hist(steps_per_day$steps, breaks=20, col="blue", main="Histogram of Total Steps Per Day", xlab="Total Steps")
dev.off()

# Mean and median of total steps per day
mean_steps <- mean(steps_per_day$steps, na.rm=TRUE)
median_steps <- median(steps_per_day$steps, na.rm=TRUE)
mean_steps
median_steps
```

## Average Daily Activity Pattern

```{r daily-pattern}
# Average steps per interval
steps_per_interval <- aggregate(steps ~ interval, activity_data, mean, na.rm = TRUE)

# Time series plot
plot(steps_per_interval$interval, steps_per_interval$steps, type="l", col="red", 
     xlab="Interval", ylab="Average Steps", main="Average Daily Activity Pattern")

# Save figure
png("figure/daily_activity_pattern.png", width=800, height=600)
plot(steps_per_interval$interval, steps_per_interval$steps, type="l", col="red", 
     xlab="Interval", ylab="Average Steps", main="Average Daily Activity Pattern")
dev.off()

# Interval with maximum steps
max_interval <- steps_per_interval[which.max(steps_per_interval$steps), ]
max_interval
```

## Imputing Missing Values

```{r impute-missing}
# Count missing values
missing_values <- sum(is.na(activity_data$steps))
missing_values

# Impute missing values using mean for that interval
activity_data_imputed <- activity_data
for (i in 1:nrow(activity_data_imputed)) {
  if (is.na(activity_data_imputed$steps[i])) {
    activity_data_imputed$steps[i] <- steps_per_interval$steps[steps_per_interval$interval == activity_data_imputed$interval[i]]
  }
}

# Recalculate total steps per day with imputed data
steps_per_day_imputed <- aggregate(steps ~ date, activity_data_imputed, sum)

# Plot histogram
hist(steps_per_day_imputed$steps, breaks=20, col="green", main="Histogram of Total Steps Per Day (Imputed)", xlab="Total Steps")

# Save figure
png("figure/total_steps_hist_imputed.png", width=800, height=600)
hist(steps_per_day_imputed$steps, breaks=20, col="green", main="Histogram of Total Steps Per Day (Imputed)", xlab="Total Steps")
dev.off()

# Mean and median after imputation
mean_steps_imputed <- mean(steps_per_day_imputed$steps)
median_steps_imputed <- median(steps_per_day_imputed$steps)
mean_steps_imputed
median_steps_imputed
```

## Weekday vs. Weekend Patterns

```{r weekday-weekend}
# Create weekday/weekend variable
activity_data_imputed$day_type <- ifelse(weekdays(activity_data_imputed$date) %in% c("Saturday", "Sunday"), "Weekend", "Weekday")

# Compute average steps per interval for weekdays and weekends
steps_weekday <- aggregate(steps ~ interval + day_type, activity_data_imputed, mean)

# Plot
p <- ggplot(steps_weekday, aes(x=interval, y=steps, color=day_type)) +
  geom_line() +
  facet_wrap(~day_type, ncol=1) +
  labs(title="Activity Patterns: Weekday vs. Weekend", x="Interval", y="Average Steps") +
  theme_minimal()

print(p)

# Save figure
ggsave("figure/weekday_weekend_plot.png", plot=p, width=8, height=5)
```

## Conclusion

- The mean and median of daily steps increased slightly after imputing missing values.
- The 5-minute interval with the maximum steps was identified.
- There is a difference in activity patterns between weekdays and weekends.

This analysis provides insights into personal activity patterns and highlights the importance of data imputation.